Parameter precision in economic growth models inspired by Ecological Economics

### 2015-10-05 Articles/Article_EcolEcon_2015/Paper0/Paper0-Resampling.Rnw
Editing intro to make generic eqns.

The Cobb-Douglas production function 
is the canonical representation of indexed economic output ($y$) 
as a function of indexed factors of production~($x_1$, $x_2$, and $x_3$) and 
an additional exogenous, time-dependent term~($A$).
This production function can be expressed as
%
\begin{equation} \label{eq:CD}
  y = \theta A \, {x_1}^{\alpha_1} \, {x_2}^{\alpha_2} \, {x_3}^{\alpha_3} \; ; \; A \equiv \mathrm{e}^{\lambda(t-t_0)},
\end{equation}
%
where 
$y \equiv Y/Y_{0}$,
$\theta$ is a scale parameter,
e is the base of the natural logarithm, 
$\lambda$ is represents the pace of technological progress,
$t$ is time measured in years,
$x_1 \equiv K/K_{0}$, 
$x_2 \equiv L/L_{0}$, 
$Y$ is economic output represented by GDP, 
$K$ is capital stock expressed in currency units, 
$L$ is labor expressed in workers or work-hours/year, and
0 subscripts indicate values at an initial year.
The parameters $\alpha_k$ and $\alpha_l$ are output elasticities
that indicate the marginal productivity of 
capital and labor in the economy, respectively.
Typically, the constraint $\alpha_k + \alpha_l = 1$ is assumed, 
indicating constant returns to scale.%
  \footnote{Dimensionless, indexed quantities are represented by lower-case symbols 
  ($y$, $k$, $l$, and later $e$), 
  and dimensional quantities are represented by upper-case symbols 
  ($Y$, $K$, $L$, and later $E$). 
  Model parameters are represented by Greek letters
  ($\theta$, $\lambda$, $\alpha_k$, $\alpha_l$, 
  later $\alpha_e$, $\delta_1$, $\delta$, $\sigma_1$, and $\sigma$).}

The capital-labor Cobb-Douglas production function shown in Equation~\ref{eq:CD}
can be augmented by energy:
%
\begin{equation} \label{eq:CDe}
  y = \theta A k^\alpha_k l^\alpha_l e^\alpha_e \; ; \; A \equiv \mathrm{e}^{\lambda(t-t_0)},
\end{equation}
%
where $e \equiv E/E_0$, and $E$ is in units of energy per time, typically TJ/year. 
The energy-augmented Cobb-Douglas production function 
is often assumed to have constant returns to scale for the three factors 
of production: $\alpha_k + \alpha_l + \alpha_e = 1$.

@@

The Constant Elasticity of Substitution~(CES) model 
without energy is given in Equation~\ref{eq:CESkl}.
%
\begin{equation} \label{eq:CESkl}
  y = \gamma \: A \: \left[\delta_1 k^{-\rho_1} 
      + (1-\delta_1)l^{-\rho_1} \right]^{-1/\rho_1}; \;
      A \equiv \mathrm{e}^{\lambda (t-t_0)} . 
\end{equation}
%
The generic CES production function
can be given as
%
\begin{equation} \label{eq:CESgeneric}
  y = \gamma \: A \: \left\{\delta \left[\delta_1 x_1^{-\rho_1} 
      + (1-\delta_1)x_2^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) x_3^{-\rho} \right\}^{-1/\rho}; 
      A \equiv \mathrm{e}^{\lambda (t-t_0)},
\end{equation}
%
where
$x_1$, $x_2$, and $x_3$ are factors of production
and permutations of capital~($k$), labor~($l$), and energy~($e$).
%
Equation~\ref{eq:CESkle} augments Equation~\ref{eq:CESkl} with energy 
using a ($kl$)$e$ nesting structure, as is typical in the literature. 
Equation~\ref{eq:CESkl} is a degenerate form of Equation~\ref{eq:CESkle} 
where $\delta \rightarrow 1$. 
%
\begin{equation} \label{eq:CESkle}
  y = \gamma \: A \: \left\{\delta \left[\delta_1 k^{-\rho_1} 
      + (1-\delta_1)l^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) e^{-\rho} \right\}^{-1/\rho}; 
      A \equiv \mathrm{e}^{\lambda (t-t_0)}.
\end{equation}
%
In the CES production function, 
$\theta$ is a fitting parameter that accounts for an atypical first year. 
The fitting parameters $\rho_1$ and $\rho$ indicate elasticities 
of substitution ($\sigma_1$ and $\sigma$).

Two other nestings of the 
factors of production---($le$)$k$ and ($ek$)$l$---are possible 
with a three-factor CES function.
%
\begin{equation} \label{eq:CESlek}
  y = \theta \: A \: \left\{\delta \left[\delta_1 l^{-\rho_1}
      + (1-\delta_1) e^{-\rho_1} \right]^{\rho/\rho_1}
      + (1-\delta) k^{-\rho} \right\}^{-1/\rho};
      A \equiv \mathrm{e}^{\lambda t}
\end{equation}
%
\begin{equation} \label{eq:CESekl}
  y = \theta \: A \: \left\{\delta \left[\delta_1 e^{-\rho_1}
      + (1-\delta_1) k^{-\rho_1} \right]^{\rho/\rho_1}
      + (1-\delta) l^{-\rho} \right\}^{-1/\rho};
      A \equiv \mathrm{e}^{\lambda t}
\end{equation}
%
Each of the CES functions (Equations~\ref{eq:CESkl} 
and~\ref{eq:CESkle}--\ref{eq:CESekl}) assumes constant returns 
to scale.

For the CES function, 
output elasticities~($\alpha$) are not constant;
rather, they vary with factors of production $k$, $l$, and $e$ over time.
For the $(kl)e$ nesting (Equation~\ref{eq:CESkle}),
output elasticities are shown in Equations~\ref{eq:alpha_k}--\ref{eq:alpha_e}.
Output elasticities for other nests can obtained by permutation of
the factors of production $k$, $l$, and $e$.
%
\begin{equation} \label{eq:alpha_k}
  \alpha_k \equiv
  \frac{
  \frac{\partial y}{\partial k} }{
  \frac{y}{k}
  } = \frac{\delta  \delta _1 k^{-\rho _1} \left[\delta _1 k^{-\rho _1}
  	+\left(1 - \delta_1 \right) 
		l^{-\rho _1}\right]{}^{\frac{\rho }{\rho _1}-1}}{\delta  
		\left[\delta _1
     k^{-\rho _1} + \left(1 - \delta _1 \right) l^{-\rho _1} \right]{}^{\frac{\rho }{\rho
     _1}} + (1 - \delta) e^{-\rho }}
\end{equation}
%
\begin{equation} \label{eq:alpha_l}
  \alpha_l \equiv 
  \frac{
  \frac{\partial y}{\partial l} }{
  \frac{y}{l}} =
  \frac{\delta  \left(1-\delta _1\right) l^{-\rho _1}
     \left[\delta _1 k^{-\rho _1}+\left(1 - \delta_1\right) l^{-\rho_1}\right]{}
	 				^{\frac{\rho }{\rho _1}-1}}
		{\delta  \left[\delta _1 k^{-\rho_1} + \left(1 - \delta_1\right) 
					l^{-\rho_1}\right]{}^{\frac{\rho }{\rho _1}}+ (1 - \delta) e^{-\rho }}
\end{equation}
%
\begin{equation} \label{eq:alpha_e}
  \alpha_e \equiv
  \frac{
  \frac{\partial y}{\partial e} }{
  \frac{y}{e}
  } =  \frac{1 - \delta}{\delta  e^{\rho } \left[\delta _1 k^{-\rho_1} 
  			+ \left(1 - \delta_1 \right) l^{-\rho _1}\right]{}
				^{\frac{\rho }{\rho_1}} + 1 - \delta}
\end{equation}
%
It can be verified that
%
\begin{equation} \label{eq:sum_betas}
  \alpha_k + \alpha_l + \alpha_e = 1
\end{equation}
%
for all nests.

>>>>

The Cobb-Douglas production function 
is the canonical representation of indexed economic output ($y$) 
as a function of indexed factors of production~($x_1$, $x_2$, and $x_3$) and 
an additional exogenous, time-dependent term~($A$).
This production function can be expressed as
%
\begin{equation} \label{eq:CD}
  y = \theta A \, {x_1}^{\alpha_1} \, {x_2}^{\alpha_2} \, {x_3}^{\alpha_3} \; ; \; A \equiv \mathrm{e}^{\lambda t},
\end{equation}
%
where 
$y \equiv Y/Y_{0}$,
$\theta$ is a scale parameter,
e is the base of the natural logarithm, 
$\lambda$ is represents the pace of technological progress,
$t$ is time measured in years,
$x_1 = k \equiv K/K_{0}$, 
$x_2 = l \equiv L/L_{0}$, 
$x_3 = e \equiv E/E_{0}$, 
$Y$ is economic output represented by GDP, 
$K$ is capital stock expressed in currency units, 
$L$ is labor expressed in workers or work-hours/year, 
$E$ is energy expressed in units of energy per time, typically TJ/year, and
0 subscripts indicate values at an initial year.
All historical data~($y$, $k$, $l$, and $e$) are indexed (by ratio)
relative to an initial year. 
Time~($t$) is indexed (by difference) relative to an initial year.
After indexing, $\theta$ is expected to be near unity,
but no other parameters change.%
  \footnote{Dimensionless, indexed quantities are represented by lower-case symbols 
  ($y$, $x$, $k$, $l$, and $e$), 
  and dimensional quantities are represented by upper-case symbols 
  ($Y$, $K$, $L$, and $E$). 
  Model parameters are represented by Greek letters
  ($\theta$, $\lambda$, $\alpha_1$, $\alpha_2$, $\alpha_3$, and
  later $\delta_1$, $\delta$, $\sigma_1$, and $\sigma$).}

The parameters $\alpha_1$, $\alpha_2$, and $\alpha_3$ are output elasticities
that indicate the marginal productivity of 
$x_1$, $x_2$, and $x_3$ in the economy, respectively.
Typically, the constraint $\alpha_1 + \alpha_2 + \alpha_3 = 1$ is assumed, 
indicating constant returns to scale.

@@

The Constant Elasticity of Substitution~(CES) model 
is given as
%
\begin{equation} \label{eq:CESgeneric}
  y = \gamma \: A \: \left\{\delta \left[\delta_1 x_1^{-\rho_1} 
      + (1-\delta_1)x_2^{-\rho_1} \right]^{\rho/\rho_1} 
      + (1-\delta) x_3^{-\rho} \right\}^{-1/\rho}; 
      A \equiv \mathrm{e}^{\lambda (t-t_0)},
\end{equation}
%
where
$x_1$, $x_2$, and $x_3$ are factors of production
and permutations of capital~($k$), labor~($l$), and energy~($e$).

For the CES function, 
output elasticities~($\alpha_1$, $\alpha_2$, and $\alpha_3$) are not constant;
rather, they vary with factors of production ($x_1$, $x_2$, and $x_3$) over time.
%
\begin{equation} \label{eq:alpha_x_1}
  \alpha_{x_1} \equiv
  \frac{ \frac{\partial y}{\partial x_1} }
          { \frac{y}{x_1} } 
  = 
  \frac{ \delta  \delta _1 x_1^{-\rho _1} \left[\delta _1 x_1^{-\rho _1} + \left(1 - \delta_1 \right) x_2^{-\rho _1}\right]
                             ^{\frac{\rho}{\rho _1}-1} }
  { \delta \left[\delta _1 x_1^{-\rho _1} + \left(1 - \delta _1 \right) x_2^{-\rho _1} \right]
                             ^{ \frac{\rho }{\rho_1} } 
                + (1 - \delta) x_3^{-\rho }}
\end{equation}
%
\begin{equation} \label{eq:alpha_x_2}
  \alpha_{x_2} \equiv \frac{ \frac{\partial y}{\partial x_2} }{\frac{y}{x_2}} =
  \frac{ \delta  \left(1-\delta _1 \right) x_2^{-\rho _1} \left[\delta _1 x_1^{-\rho _1}+\left(1 - \delta_1\right) x_2^{-\rho_1}\right]
	 				^{\frac{\rho }{\rho _1}-1} }
		{\delta  \left[\delta _1 x_1^{-\rho_1} + \left(1 - \delta_1\right) x_2
                                        ^{-\rho_1}\right]^{\frac{\rho }{\rho _1}}+ (1 - \delta) {x_3}^{-\rho }}
\end{equation}
%
\begin{equation} \label{eq:alpha_x_3}
  \alpha_{x_3} \equiv \frac{ \frac{\partial y}{\partial x_3} }{ \frac{y}{x_3} } =  
      \frac{1 - \delta}{\delta  x_3^{\rho } \left[\delta _1 x_1^{-\rho_1} + \left(1 - \delta_1 \right) x_2^{-\rho _1}\right]
				^{\frac{\rho }{\rho_1}} + 1 - \delta}
\end{equation}
%
It can be verified that
%
\begin{equation} \label{eq:sum_betas}
  \alpha_{x_1} + \alpha_{x_2} + \alpha_{x_3} = 1
\end{equation}
%
for all nests.
